---
title: "Untitled"
author: "Ali"
date: "01/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## class March 01, 2021

- keeps working on Hodgkin-Huxley model (HH) and Fitzhgo-Nagumo model
- phase plans, computational work
- intro to Chaos

- look at H-H model, and look at the V-m phase plane
  - finding the unstable manifold is hard
  - it is helpfull

```{r packages}
library("reticulate") ## use_python("/usr/bin/python3")
```

```{r HHmodel}
parms0 <- c(g_bar_Na=120,g_bar_K=36,g_L=0.3, v_Na=-115, v_K=12, v_L=-10.5989,
C=1,I=0)
alpha <- function(v,type) {
switch(type,
m=0.1*(v+25)/(exp((v+25)/10) -1),
h=0.07*exp(v/20),
n=0.01*(v+10)/(exp((v+10)/10) -1)
)
}
beta <- function(v,type) {
switch(type,
m=4*exp(v/18),
h=1/exp((v+30)/10 + 1),
n=0.125*exp(v/80)
)
}
HHgrad <- function(t,y,parms) {
g <- with(as.list(c(y,parms)),
c(v=-1/C*(-I + g_bar_Na*m^3*h*(v-v_Na) +
g_bar_K*n^4*(v-v_K) +
g_L*(v-v_L)),
n=alpha(v,"n")*(1-n) - beta(v,"n")*n,
m=alpha(v,"m")*(1-m) - beta(v,"m")*m,
h=alpha(v,"h")*(1-h) - beta(v,"h")*h)
)
list(g)
}
y0 <- c(v=0,n=0.3,m=0.05,h=0.6)
HHgrad(0,y0,parms0)

plot_hh <- function(h) {
op <- par(mfrow=c(1,2),las=1,bty="l")
plot(h[,1],h[,2], type="l",xlab="time",ylab="V")
cvec <- c(1,2,4) ## colours
matplot(h[,1],h[,3:5], type="l",lty=1,xlab="time",ylab="", col=cvec)
legend("topright",legend=c("n","m","h"),lty=1,col=cvec)
}

library(deSolve)
res0 <- ode(y=y0,times=seq(0,60,by=0.05), func=HHgrad, parms=parms0)
plot_hh(res0)
## so far I am running the HH model numerically, the bifurcation diagram is more systematic

## bifurcation diagram...
library(phaseR) ## convenient phase plane analysis 
res0 <- ode(y=y0,times=seq(0,60,by=0.05), func=HHgrad, parms=parms0)
tail(res0,1) ## the last element of solution
HHeq <- res0[nrow(res0),c("v","n","m","h")] ## Lazy way of finding equilibria

## Now I want to redefine the dynamics in the V-m plane
HHgrad2d <- function(t, y, parms) {
full_y <- c(v=y[["v"]],m=y[["m"]],n=HHeq[["n"]],h=HHeq[["h"]]) ## n and h are set at the equil value
list(HHgrad(t,full_y,parms)[[1]][c("v","m")])## call HH grad model with n & h set at equi, rerun and pull out the v and m values and put it back to the list
}
HHgrad2d(0,y0[c("v","m")], parms0)


phasePlaneAnalysis(HHgrad2d,xlim=c(-120,10),
parameters=parms0,
state.names=c("v","m"),
ylim=c(0,1))
```


```{r python_neuro}
library(reticulate)

```

## class March 03, 2021 

- paper was discussed; Dynamical phases of the Hindmarsh-Rose neuronal model: Studies of the transition from bursting to spiking chaos
Chaos 17, 043128 (2007); (https://doi.org/10.1063/1.2818153)
Giacomo Innocenti et al

## class March 04, 2021 

more talking about chaos









