---
title: "Untitled"
author: "Ali"
date: "10/02/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## class Feb 10 2021

Linda Allen 2007 paper

- History;
  - SIR model by Kermac _Mc in 19s
  - Read-Frost model 1927 1928 report of classes, in 1951 Read made a TV show by John Hopkings youtube, in 1952 Hellen Abby his student pulished, in 1976 someone published the 1928 original work.
    - SIR model is cts in time and cts populations, deterministic, it has overlaping generations (if I can be infected as my grand childe)
    - RF model is disc time, non-overlaping
    


```{r Feb10}
## never == compare numbers wi
## To run binomial reed-Frost process
## Note that in Reed-Frost model, the I(t) is prevelence and at the same time it is incidence since the R-F model is a special case.
nt <- 20 ## time steps
y0 <- c(S=99,I=1) ## starting values

## a realization of a stochasticity proc is a single draw
S <- y0[["S"]]
I <- y0[["I"]]
p_i <- 0.02 ## so I am going to get R0=2
res <-numeric(nt) ## vector for results
res[1] <- I

for(t in 2:nt){
  new_inf <- rbinom(1,size = S,prob = 1-(1-p_i)^I)
  I <- new_inf
  S <- S-I
  if(I==0) break
  res[t] <- I
}

plot(res, type="s") ## stairsteps
res

s=99
p_i=0.02
I=1
## prob of no infection
0.98^99


## this was 1 realization, now I want to turn into a loop and a function
set.seed(101)
nt <- 20 ## time steps
y0 <- c(S=99,I=1) ## starting values
p_i <- 0.02 ## so I am going to get R0=2

rf <- function(nt,y0,p_i) {
res <- numeric(nt)
S <- y0[["S"]]
I <- res[1] <- y0[["I"]]
for (t in 2:nt) {
I <- rbinom(1, size=S, prob=1-(1-p_i)^I)
S <- S-I
if (I==0) break
res[t] <- I
}
return(res)
}
r0 <- rf(20, y0=c(S=99,I=1), p_i=0.02)
r1 <- replicate(300, rf(20, y0=c(S=99,I=1), p_i=0.02)) ## doing the same thing some specified time here 300 times
par(las=1,bty="l")
matplot(r1,type="l",col=adjustcolor("black",alpha.f=0.2),lty=1) ## plot the columns of a matrix
lines(rowMeans(r1),col=2,lwd=3) ## ave # of infections
plot(table(colSums(r1))) ## final sizes of the epidemic (note in R-F model, the prevelence is the same as incidence )
## why bimodal? either it dies out immidemately of there will be an epidemic
## what happen if we make the pop size bigger? less variability in in the I_t


## Note; if you wonder what the heck is going on with table(), try this:
# plot(table(c(1,1,1,2,3)))


```
The R-F model is a general building block of any stochastic discrete time model.

## Feb 22

- we left on Glispie algo, cts time, discrete state, stoch model
  - computing the Poisson proc
  - it is very slow, scales with the pop size
- TAU-leaping (discreatize of Glispie alg, let's set delta t to fix time, where we can have a section of time that not many changes will happen) approx Poisson or binomial how about the negative binomial?
  - negative binomial; the poisson dist assumes that an event happens prob in reality there are hetherogenathy. you can get by Bernoulli coin flipping proccess, how many trials before k successes?
  also can be thinked of as a overdispersed Poisson lambda_i ~ Gamma(mean,shape)
  X_i ~ Poisson()

```{r Feb22}
plot(table(rpois(1000,lambda=4)))

plot(table(rnbinom(1000,mu=4,size=1)))
## Poisson distribution with an underlying Gamma-dist heterogeneity: compound distribution


## Glisspie algorithm
library(GillespieSSA2)

```

- stoch ODE; cts time, stochastic, cts states
  - Poisson proc -> central limit theor, changes in pop size will be Gaussian
  - cts time-cts-state stoch proc is a weired proccess (because it is jummping around at any t so non differentialable, see Linda Allen)
  - dX/dt=deterministic+stochastic
    - deter: f(X(t)) expected change
    - stoch: G(X(t)) dW(t) expected variance of changes in covariance matrix

## Feb 24
 models of nerves excitation; mainly Hodkin-Huxley model
 - limit cycles
 - fast-slow sys
  Hodkin-Huxley model; what happen in a particular patch that gives propagate signal to pass



```{r Feb24}

parms0 <- c(g_bar_Na=120,g_bar_K=36,g_L=0.3, v_Na=-115, v_K=12, v_L=-10.5989,
C=1,I=0)
alpha <- function(v,type) {
switch(type,
m=0.1*(v+25)/(exp((v+25)/10) -1),
h=0.07*exp(v/20),
n=0.01*(v+10)/(exp((v+10)/10) -1)
)
}
beta <- function(v,type) {
switch(type,
m=4*exp(v/18),
h=1/exp((v+30)/10 + 1),
n=0.125*exp(v/80)
)
}
HHgrad <- function(t,y,parms) {
g <- with(as.list(c(y,parms)),
c(v=-1/C*(-I + g_bar_Na*m^3*h*(v-v_Na) +
g_bar_K*n^4*(v-v_K) +
g_L*(v-v_L)),
n=alpha(v,"n")*(1-n) - beta(v,"n")*n,
m=alpha(v,"m")*(1-m) - beta(v,"m")*m,
h=alpha(v,"h")*(1-h) - beta(v,"h")*h)
)
list(g)
}
y0 <- c(v=0,n=0.3,m=0.05,h=0.6)

HHgrad(0,y0,parms0)


plot_hh <- function(h) {
op <- par(mfrow=c(1,2),las=1,bty="l")
plot(h[,1],h[,2], type="l",xlab="time",ylab="V")
cvec <- c(1,2,4) ## colours
matplot(h[,1],h[,3:5], type="l",lty=1,xlab="time",ylab="", col=cvec)
legend("topright",legend=c("n","m","h"),lty=1,col=cvec)
}
library(deSolve)
res <- ode(y=y0,times=seq(0,60,by=0.05), func=HHgrad, parms=parms0)
plot_hh(res)



```





