---
title: "Untitled"
author: "Ali"
date: "10/02/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## class Feb 10 2021

Linda Allen 2007 paper

- History;
  - SIR model by Kermac _Mc in 19s
  - Read-Frost model 1927 1928 report of classes, in 1951 Read made a TV show by John Hopkings youtube, in 1952 Hellen Abby his student pulished, in 1976 someone published the 1928 original work.
    - SIR model is cts in time and cts populations, deterministic, it has overlaping generations (if I can be infected as my grand childe)
    - RF model is disc time, non-overlaping
    


```{r Feb10}
## never == compare numbers wi
## To run binomial reed-Frost process
## Note that in Reed-Frost model, the I(t) is prevelence and at the same time it is incidence since the R-F model is a special case.
nt <- 20 ## time steps
y0 <- c(S=99,I=1) ## starting values

## a realization of a stochasticity proc is a single draw
S <- y0[["S"]]
I <- y0[["I"]]
p_i <- 0.02 ## so I am going to get R0=2
res <-numeric(nt) ## vector for results
res[1] <- I

for(t in 2:nt){
  new_inf <- rbinom(1,size = S,prob = 1-(1-p_i)^I)
  I <- new_inf
  S <- S-I
  if(I==0) break
  res[t] <- I
}

plot(res, type="s") ## stairsteps
res

s=99
p_i=0.02
I=1
## prob of no infection
0.98^99


## this was 1 realization, now I want to turn into a loop and a function
set.seed(101)
nt <- 20 ## time steps
y0 <- c(S=99,I=1) ## starting values
p_i <- 0.02 ## so I am going to get R0=2

rf <- function(nt,y0,p_i) {
res <- numeric(nt)
S <- y0[["S"]]
I <- res[1] <- y0[["I"]]
for (t in 2:nt) {
I <- rbinom(1, size=S, prob=1-(1-p_i)^I)
S <- S-I
if (I==0) break
res[t] <- I
}
return(res)
}
r0 <- rf(20, y0=c(S=99,I=1), p_i=0.02)
r1 <- replicate(300, rf(20, y0=c(S=99,I=1), p_i=0.02)) ## doing the same thing some specified time here 300 times
par(las=1,bty="l")
matplot(r1,type="l",col=adjustcolor("black",alpha.f=0.2),lty=1) ## plot the columns of a matrix
lines(rowMeans(r1),col=2,lwd=3) ## ave # of infections
plot(table(colSums(r1))) ## final sizes of the epidemic (note in R-F model, the prevelence is the same as incidence )
## why bimodal? either it dies out immidemately of there will be an epidemic
## what happen if we make the pop size bigger? less variability in in the I_t


## Note; if you wonder what the heck is going on with table(), try this:
# plot(table(c(1,1,1,2,3)))


```
The R-F model is a general building block of any stochastic discrete time model.


```{r Feb11}


```



